A Machine Learning Approach to MLB Catcher Framing

### 2021-12-09 Slides.Rmd
Boosting again

- Baseball catchers can influence the call of a ball or strike on how they catch the ball
- Some catchers are better than others at this skill
- Baseball teams are aware of this and are acquiring players good at this skill to win more games
- We want to quantify the best catcher's at framing for the 2021 season
- There are several factors that influence whether a pitch will be a strike or ball
- Catchers getting more strikes translates to more outs, and fewer points for the opposing team

@@

- 3 Fold Cross Validation on Random Forest  
  - Tested on nodesizes of 10, 30, 50, 70, 90 | 500 trees
  - Compared Out-Of-Bag Error 
```{r, echo=FALSE}
load("Data/pitchData.Rda")
load("Model-Outputs/random-forest.Rda")

@@

```{r, echo=FALSE}
load("Model-Outputs/boosting.Rda")

@@

>>>>

- Baseball catchers can influence the call of a ball or strike on how they catch the ball
- Some catchers are better than others at this skill
- Baseball teams are aware of this and are acquiring players good at this skill to win more games
- We want to quantify the best catcher's at framing for the 2021 season
- There are several factors that influence whether a pitch will be a strike or ball
- Catchers getting more strikes translates to more outs, and fewer runs for the opposing team

@@

- Tested several Random Forest models
  - Nodesize: 10, 30, 50, 70, 90 | # of Trees: 500
  - Compared Out-Of-Bag Error 
```{r, echo=FALSE, fig.width=7, fig.height=3}
load("Data/pitchData.Rda")
load("Model-Outputs/random-forest.Rda")

@@

- 3 Fold Cross Validation on Boosting
  - Interaction depth: 6, 11, 16 
  - Shrinkage: 0.1, 0.01, 0.001 
  - Number of Trees: 500
  - Compared CV accuracy
- 27 models took over 5 hours to run

```{r, echo=FALSE, fig.width=6, fig.height=2.5}
load("Model-Outputs/boosting.Rda")

@@

## Boosting Part 2

- Only tuned shrinkage and interaction depth
- Now tune trees with interaction depth of 16 and shrinkage of 0.1

```{r}
load("Model-Outputs/boosting2.Rda")
final_df_gbm %>%
  group_by(trees) %>%
  summarize(accuracy = mean(accuracy)) %>%
  arrange(desc(accuracy)) %>%
  ggplot(aes(x = trees, y = accuracy)) +
  geom_line(color = "goldenrod") +
  theme_minimal() +
  labs(x = "Number of Trees",
       y = "Accuracy",
       title = "Boosting Accuracy by # of Trees")
```
